My title

### 2024-11-03 paper/paper.qmd
Diagnostics for model update

# Results
To assess model reliability, we examined several key diagnostics. Convergence metrics, such as Rhat values, were close to 1 for all parameters, indicating strong convergence. Additionally, the effective sample size (n_eff) was high across parameters, suggesting low autocorrelation and contributing to model stability. For validation, we conducted out-of-sample testing and calculated the Root Mean Square Error (RMSE) to assess predictive accuracy. This test data, which includes polls not used in training, provides an unbiased estimate of model performance.

@@

Table @tbl-model-coefficients presents the estimated coefficients for the predictors in our GLM model. These coefficients fit into the GLM equation, allowing us to interpret the impact of each predictor on Harris’s predicted vote percentage. Positive values indicate a higher predicted percentage, while negative values indicate a decrease. Key predictors, such as specific pollsters and population types, show distinct effects on the forecasted outcomes. This table is produced using kable from knitr for enhanced readability.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-coefficient-estimates
#| fig-cap: Coefficient Estimates for Predictors
coefficients %>%
  ggplot(aes(estimate, term)) + 
  geom_point() + 
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) + 
  labs(title = "Coefficient Estimates", x = "Estimate", y = "Predictor") +
  theme(axis.text.y = element_text(size = 7))
```

Figure @fig-coefficient-estimates visually represents the model coefficients, with error bars indicating the confidence interval for each estimate. Positive coefficients suggest that specific pollsters or demographic characteristics increase Harris’s predicted vote share. For example, the 'CES / YouGov' pollster shows a positive impact, while 'Siena/NYT' has a slight negative effect. These error bars help contextualize the reliability of each predictor.

@@

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

#pp_check(first_model) +
#  theme_classic() +
 # theme(legend.position = "bottom")

#posterior_vs_prior(first_model) +
#  theme_minimal() +
#  scale_color_brewer(palette = "Set1") +
 # theme(legend.position = "bottom") +
 # coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

#plot(first_model, "trace")

#plot(first_model, "rhat")
```

>>>>

# Results
To assess model reliability, we examined several key diagnostics. Convergence metrics, such as Rhat values, were close to 1 for all parameters, indicating strong convergence. Additionally, the effective sample size (n_eff) was high across parameters, suggesting low autocorrelation and contributing to model stability. For validation, we conducted out-of-sample testing and calculated the Root Mean Square Error (RMSE) to assess predictive accuracy. This test data, which includes polls not used in training, provides an unbiased estimate of model performance. See more details of model diagnostic here @sec-model-details-diagnostics.

@@

@tbl-model-coefficients presents the estimated coefficients for the predictors in our GLM model. These coefficients fit into the GLM equation, allowing us to interpret the impact of each predictor on Harris’s predicted vote percentage. Positive values indicate a higher predicted percentage, while negative values indicate a decrease. Key predictors, such as specific pollsters and population types, show distinct effects on the forecasted outcomes.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-coefficient-estimates
#| fig-cap: Coefficient Estimates for Predictors
coefficients %>%
  ggplot(aes(estimate, term)) + 
  geom_point() + 
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high)) + 
  labs(x = "Estimate", y = "Predictor") +
  theme(axis.text.y = element_text(size = 7))
```

@fig-coefficient-estimates represents the model coefficients, with error bars indicating the confidence interval for each estimate. Positive coefficients suggest that specific pollsters or demographic characteristics increase Harris’s predicted vote share. For example, the 'CES / YouGov' pollster shows a positive impact, while 'Siena/NYT' has a slight negative effect. These error bars help contextualize the reliability of each predictor.

@@

# Diagnostics for model {#sec-model-details-diagnostics}

@fig-ppcheck compares observed data (dark line) with replicated posterior predictions (lighter lines). The close alignment suggests that the model accurately captures the data's central tendency and variability. @fig-convergencecheckTrace and @fig-convergencecheckRhat show that the sampling algorithm used, the Markov chain Monte Carlo (MCMC) algorithm, did not run into issues as the posterior distribution for the model was created. Using the checks presented by @citetellingstorieswithdata, both graphs do not show anything abnormal since he trace plots in @fig-convergencecheckTrace display substantial horizontal fluctuation across chains, indicating good mixing, while the Rhat values in @fig-convergencecheckRhat are close to 1 and well below 1.1, further supporting convergence.
```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-ppcheck
#| fig-cap: "Posterior Predictive Check: Comparison of Observed and Replicated Data"
pp_check(harris_model) + theme_minimal()
```

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-convergencecheckTrace
#| fig-cap: "Checking the convergence of the MCMC algorithm - Trace"

plot(harris_model, "trace") + theme_minimal()

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-convergencecheckRhat
#| fig-cap: "Checking the convergence of the MCMC algorithm - Rhat"

plot(harris_model, "rhat") + theme_minimal()
